---
title: "Practical Machine Learning Project"
author: "inyigolopez"
date: "December, 2015"
output: html_document
---

## *Background*

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, **your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.** More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

## *Data*

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

## *Goal*

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

## *Reading and Cleaning data*

```{r echo=FALSE}
library(lattice)
library(ggplot2)
library(caret)
library(kernlab)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
```

### Reading

First of all we need read the data. I downloaded the data to my workspace and put empty cells as NA data.

```{r}

training <- read.csv(file = 'pml-training.csv', row.names = 1, na.strings=c("NA","#DIV/0!",""))
test <- read.csv(file = 'pml-testing.csv',row.names = 1, na.strings=c("NA","#DIV/0!",""))

dim(training)

```

### Cleaning 

I need good variables to make my predictions and my training data have 159 variables but the are a lot of NA data. I select all variables that have minus than 10% of NA data and are significant in the variance measure.

```{r}
nsv <- nearZeroVar(training, saveMetrics = T)
training <- training[, !nsv$nzv]

training <- training[, colSums(is.na(training)) < nrow(training)*0.1]

```

...and i drop the first variable, X, because is just the number of row, not a variable:

```{r}
training <- training[,2:dim(training)[2]]
```



### *Split training data into training and cross validation*

I select 70% of data for training and 30% for cross-validation

```{r}
set.seed(23425331)
inTrain <- createDataPartition(y=training$classe, p=0.7, list=FALSE)

training <- training[inTrain,]
crossValidation <- training[-inTrain,]

```

## Training Random Forest model

Now I train our model with all variables to predict 'classe' variable, using Random Forest because is an algorithm with a big accuracy:
```{r}
mfControl <- trainControl(method="cv", number=3, verboseIter=F)
modFit <- train(classe ~ ., data=training, method="rf", trControl=mfControl)
```

## Evaluate the model over cross validation data

And now, we use our **cross validation** data to see the prediction of our model over this dataset and the accuracy of our model:
```{r}
pred_CV_RF <- predict(modFit, crossValidation)
confusionMatrix(pred_CV_RF,crossValidation$classe)
```

We see the acuraccy of our model is very high and we can use our model to predict 'classe' over **test data** :

## Predict over test data

```{r}
pred_TEST_RF <- predict(modFit, test)
pred_TEST_RF
```

## Generating Files for submit

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(pred_TEST_RF)
```








